Subtle Gestures with the J!NS MEME

### 2021-08-09 Paper/paper.tex
Formulate EOG/subtle subsection

% Assigned to FB
\section{Related Work}
[Insert an introduction here]

\subsection{Face Interaction}
Detecting finger tapping on the body using acoustic sensing \cite{10.1145/1753326.1753394}.
Optical tracking system with infrared markers on fingers \cite{10.1145/2556288.2556984}.
Exploring touch interaction by placing electrodes around the ear \cite{10.1145/2468356.2468592}.
Exploring tooth click gestures \cite{10.1145/2935334.2935389}.
Eyelid gestures \cite{10.5555/2788890.2788938}.
Finger sensing IMUs \cite{10.1145/2858036.2858466}.
Hand sensing IMUs \cite{10.1145/2984511.2984582}.
Gesture recognition using wrist band \cite{10.1145/3274783.3274854}.
Tracking tongue movements with touch sensors \cite{10.1145/3311823.3311831}.
\subsection{Applications for Wearable Sensing}
Tracking eating episodes \cite{10.1145/3130902}.
Sensors in shoes for tracking exercises \cite{10.1145/3174910.3174938}.
Using hand gestures to control TV with wrist band \cite{10.1145/2641248.2641359}.
Jump performance analysis \cite{10.1145/2753509.2753512}.
Monitoring drinking with wrist-worn inertial sensors \cite{10.1145/3267242.3267253}.
\subsection{EOG Sensing and Subtle Interaction}
EOG attached to over-ear head phones \cite{10.1145/1125451.1125655}.
EOG attached to in-ear head phones \cite{10.1145/2493988.2494329}.
Demonstration of EOG sensors for eye-tracking \cite{10.1145/1520340.1520468}.
Commercial EOG glasses for activity recognition, passively \cite{10.1145/2638728.2638795}.
KissGlass for passively sensing context, not for deliberate subtle interactions \cite{10.1145/3384657.3384801}.
Defining subtlety \cite{10.1145/3290605.3300648}. 
Itchy Nose \cite{10.1145/3123021.3123060}.
% Should this one even be in the Related Work category or just mentioned in Analysis?
Itchy Nose tool for automating personalization \cite{10.1145/3174910.3174953}.

% Assigned to FB
\section{Analysis}
What was our workflow?

@@

What data do we collect, how and why?

>>>>

% Assigned to FB
\section{Related Work}
Our work touches on multiple areas of research, including wearable sensing, gesture recognition,
and subtle interaction.

\subsection{Face Interaction}
Detecting finger tapping on the body using acoustic sensing \cite{10.1145/1753326.1753394}.
Optical tracking system with infrared markers on fingers \cite{10.1145/2556288.2556984}.
Exploring touch interaction by placing electrodes around the ear \cite{10.1145/2468356.2468592}.
Eyelid gestures \cite{10.5555/2788890.2788938}.
Finger sensing IMUs \cite{10.1145/2858036.2858466}.
Hand sensing IMUs \cite{10.1145/2984511.2984582}.
Gesture recognition using wrist band \cite{10.1145/3274783.3274854}.
% Applications for Wearable Sensing
Tracking eating episodes \cite{10.1145/3130902}.
Sensors in shoes for tracking exercises \cite{10.1145/3174910.3174938}.
Using hand gestures to control TV with wrist band \cite{10.1145/2641248.2641359}.
Jump performance analysis \cite{10.1145/2753509.2753512}.
Monitoring drinking with wrist-worn inertial sensors \cite{10.1145/3267242.3267253}.

\subsection{EOG Sensing and Subtle Interaction}
While the aforementioned work utilized mainly IMUs, touch sensors or optical systems,
Electrooculography (EOG) has become a useful tool for eye-tracking in recent years, too. 
In 2006, Manabe et al. attached EOG sensors to over-ear \cite{10.1145/1125451.1125655}
and in 2013, to in-ear head phones \cite{10.1145/2493988.2494329} to detect eye gestures.
Bulling et al. first demonstrated efficient real-time eye-movement recognition
with EOG goggles in 2009, which they custom-made \cite{10.1145/1520340.1520468}.
Once commercial EOG glasses became available in 2014, Ichimaru et al. used them for
activity recognition \cite{10.1145/2638728.2638795}. They used the J!ns Meme
smartglasses to distinguish between typing, reading, eating and talking.
As for more recent work, Li et al. used the J!ns Meme to infer the physical and social
context of the wearer by recognizing various greeting gestures involving kissing
\cite{10.1145/3384657.3384801}.

In contrast to the work discussed, we focused on subtle interaction. Pohl et al. define
four types of subtle interaction: (1) signifying feedback that is non-intrusive to the
user, (2) hiding interaction from others and potentially deceiving them, (3) employing
less effort for input and generally doing less, and (4) nudging users
\cite{10.1145/3290605.3300648}. Our work aims at type (2) social subtlety, like e.g.
Ashbrook et al. did with their tooth-click gesture recognition interface
\cite{10.1145/2935334.2935389}. With a similar goal, Li et al. developed TongueBoard,
a retainer form-factor device for recognizing silent speech \cite{10.1145/3311823.3311831}.

The paper \textit{Itchy Nose} by Lee et al. is the most closely related work to ours,
as it inspired our efforts \cite{10.1145/3123021.3123060}. They used the J!ns Meme
smartglasses to detect multiple gestures that consisted of rubbing or pushing their
noses, which can also be detected by the glasses' EOG sensors.

% Assigned to FB
\section{Analysis}
Explaining the J!ns Meme.

What was our workflow?

@@

What data do we collect, how and why?
Itchy Nose tool for automating personalization \cite{10.1145/3174910.3174953}.
